---
title: Overview
---

Addresses represent a physical place. They're meant to be interpreted by people and help guide navigation of the built environment. Addresses represent a geographical place but lack geographic data.

The package `{arcgisgeocode}` enables you to search for an address (geocode), reverse geocode, find candidate matches, get suggestions, and batch geocode. Geocoding is the process of converting text to an address and a location.

- **Address geocoding**, also known as forward geocoding, is the process of converting text for an address to a complete address with a location.
- **Place geocoding** is the process of searching for addresses for businesses, administrative locations, and geographic features.
- **Reverse geocoding** is the process of converting a point to an address or place.
- **Batch geocoding**, also known as bulk geocoding, is the process of converting a list of addresses or place names to a set of complete addresses with locations.

## Licensing considerations

Many features of the ArcGIS World Geocoder are provided for free such as forward geocoding, reverse geocoding, and place search. However, **storing results is not free**. Additionally, the bulk geocoding functionality requires a developer account or available credits. 

In order to store results, each function has an argument `for_storage` which should be set to `TRUE` if you intend to store the results. 

To learn more about free and paid geocoding operations refer to the [storage parameter documentation](https://developers.arcgis.com/documentation/mapping-apis-and-services/geocoding/services/geocoding-service/#storage-parameter).

| Function | Description | Free |
| -------- | ----------- | ---- |
| `find_address_candidates()` | Finds up to 50 location candidates based on a provided address. _This function is vectorized_ to work with many addresses at a time. | ✅ |
| `reverse_geocode()` | Returns an address based on the provided coordinate. _This function is vectorized_ to work with many locations at a time. | ✅ |
| `suggest_places()` | Returns possible POI information based on a location and a search phrase. This function is not vectorized. | ✅ |
| `geocoded_addresses()` | Bulk geocodes addresses returning a single location per address. Use this for highly performant and scalable address geocoding. | ❌ |


## Get started

To start geocoding with the R-ArcGIS Bridge, install the R package from CRAN. 

```{r include = FALSE}
library(pillar)
```
```r
# install from CRAN
install.packages("arcgisgeocode")
```

```{r}
# Load the library
library(arcgisgeocode)
```

### Geocode an address

Perform single address geocoding using the `find_address_candidates()` function. Limit the number of results using the `max_locations` argument. 

```{r}
loc <- find_address_candidates(
    "501 Edgewood Ave SE, Atlanta, GA 30312", max_locations = 1
)

loc[, 1:8]
```

### Reverse geocode 

From a location, find its corresponding address using `reverse_geocode()`. 

```{r message=FALSE}
reverse_geocode(c(-84.371, 33.753))
```
