---
title: Find routes and directions
uid: finding-routes
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

Routing, also known as simple routing, is the process of finding the best path from an origin to a destination for an individual or a single vehicle. It accounts for network data such as speed limit, number of lanes, and time of day, and it can incorporate real-time conditions such as traffic, accidents, and other barriers.

A route can be **simple** or **optimized**. A simple route is the shortest path between an ordered set of stops. An optimized route reorders the stops to find the most efficient path between them. `find_routes()` provides both, and it can also return turn-by-turn directions for the trip.

## Step 1. Authenticate

Routing services require an authenticated token. Set one for your session using `{arcgisutils}`.

```{r}
library(sf)
library(arcgisutils)
library(arcgisrouting)

set_arc_token(auth_user())
```

## Step 2. Define stops

`find_routes()` takes `stops` as an `sf` object of point geometries. Create three stops in San Francisco using `st_sf()`.

```{r}
stops <- st_sf(
  name = c("Ferry Building", "Coit Tower", "Painted Ladies"),
  geometry = st_sfc(
    st_point(c(-122.3933, 37.7955)),
    st_point(c(-122.4058, 37.8024)),
    st_point(c(-122.4329, 37.7762)),
    crs = 4326
  )
)

stops
```

## Step 3. Find the route

Pass the stops to `find_routes()`. By default the route visits the stops in the order provided and includes both the route geometry and turn-by-turn directions.

```{r}
route <- find_routes(stops)
route$routes
```

The `directions` element contains the turn-by-turn directions for the trip.

```{r}
route$directions |> 
  tidyr::unnest(directions) |> 
  dplyr::select(time, length, text)
```

## Optimizing the stop order

Set `find_best_sequence = TRUE` to reorder the stops and find the most efficient route. Use `preserve_first_stop` and `preserve_last_stop` to keep the trip's starting or ending point fixed while the stops in between are reordered.

```{r}
optimized <- find_routes(
  stops,
  find_best_sequence = TRUE,
  preserve_first_stop = TRUE,
  preserve_last_stop = TRUE
)

optimized$routes
```

## Choosing a travel mode

`find_routes()` defaults to a standard driving travel mode. Use `travel_mode` to route by walking or trucking instead, or to apply a travel mode defined by your organization. See `get_travel_modes()` to discover the travel modes available to you.

```{r}
modes <- get_travel_modes()
modes
```

## Larger routing jobs

`find_routes()` makes a direct request, which suits shorter transactions with fewer than 150 stops that complete in under five minutes. For longer transactions, up to 10,000 stops completing in under 60 minutes, use `find_routes_job()` instead, which creates a geoprocessing job.
