---
title: Last mile delivery
uid: last-mile-delivery
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

Last mile delivery is a type of vehicle routing problem tailored to the unique challenges of final deliveries in dense urban environments. Rather than simply minimizing overall cost, it produces geographically clustered routes so drivers can visit nearby stops without unnecessary detours, minimizing operating cost for the fleet.

Last mile delivery only supports asynchronous job requests. `last_mile_delivery()` takes the same core inputs as `route_vehicles()`, `orders`, `depots`, and `routes`, along with delivery-specific inputs such as `zones` and specialty requirements.

## Step 1. Authenticate

```{r}
library(sf)
library(arcgisutils)
library(arcgisrouting)

set_arc_token(auth_user())
```

## Step 2. Define orders, depots, and routes

Each order can specify a time window during which the delivery must occur, and how much of the vehicle's capacity it requires.

```{r}
orders <- st_as_sf(
  data.frame(
    name = c("Order 1", "Order 2"),
    service_time = c(5, 5),
    time_window_start = as.POSIXct(
      c(NA, 1706860800),
      origin = "1970-01-01",
      tz = "UTC"
    ),
    time_window_end = as.POSIXct(
      c(1706868000, 1706868000),
      origin = "1970-01-01",
      tz = "UTC"
    ),
    max_violation_time = c(0, 30),
    delivery_quantity_1 = c(2000, 1500),
    delivery_quantity_2 = c(100, 75),
    x = c(-117, -117.5),
    y = c(34, 34.5)
  ),
  coords = c("x", "y"),
  crs = 4326
)

depots <- st_as_sf(
  data.frame(name = "Depot 1", x = -117.2, y = 34.2),
  coords = c("x", "y"),
  crs = 4326
)

routes <- data.frame(
  name = c("Truck 1", "Truck 2"),
  start_depot_name = c("Depot 1", "Depot 1"),
  end_depot_name = c("Depot 1", "Depot 1"),
  earliest_start_time = c("6:00:00", "6:00:00"),
  capacity_1 = c(40000, 30000),
  capacity_2 = c(2000, 2500),
  cost_per_unit_time = c(0.5, 0.5),
  cost_per_unit_distance = c(1.5, 1.5)
)
```

## Step 3. Submit the job

```{r}
job <- last_mile_delivery(
  orders = orders,
  depots = depots,
  routes = routes,
  earliest_route_start_date = as.Date("2024-02-02"),
  max_route_total_time = 480,
  time_units = "minutes",
  route_shape = "true_shape_with_measures",
  populate_directions = TRUE
)

job$start()
job$await()
```

## Specialties and zones

Some orders require a vehicle with a particular capability, such as refrigeration or hazardous materials handling. Use `order_specialties` and `route_specialties` to match orders to routes that support them.

```{r}
order_specialties <- data.frame(
  order_name = c("Order 1", "Order 2"),
  specialty_name = c("Refrigerated", "Hazmat")
)

route_specialties <- data.frame(
  route_name = c("Truck 1", "Truck 2"),
  specialty_name = c("Refrigerated", "Hazmat")
)
```

Use `zones` to assign each route to a work territory, keeping its deliveries within a defined polygon.

```{r}
zone1 <- st_polygon(list(rbind(
  c(-97.0634, 32.8442),
  c(-97.0554, 32.84),
  c(-97.0558, 32.8327),
  c(-97.0638, 32.83),
  c(-97.0634, 32.8442)
)))

zone2 <- st_multipolygon(list(
  list(rbind(
    c(-97.0803, 32.8235),
    c(-97.0776, 32.8277),
    c(-97.074, 32.8254),
    c(-97.0767, 32.8227),
    c(-97.0803, 32.8235)
  )),
  list(rbind(
    c(-97.0871, 32.8311),
    c(-97.0831, 32.8292),
    c(-97.0853, 32.8259),
    c(-97.0892, 32.8279),
    c(-97.0871, 32.8311)
  ))
))

zones <- st_cast(
  st_sf(
    name = c("Zone 1", "Zone 2"),
    geometry = st_sfc(zone1, zone2, crs = 4326)
  ),
  "MULTIPOLYGON"
)

job <- last_mile_delivery(
  orders = orders,
  depots = depots,
  routes = routes,
  order_specialties = order_specialties,
  route_specialties = route_specialties,
  zones = zones,
  earliest_route_start_date = as.Date("2024-02-02"),
  max_route_total_time = 480,
  sequence_gap = 3,
  time_units = "minutes",
  route_shape = "true_shape_with_measures",
  populate_directions = TRUE
)

job$start()
job$await()
```
