---
title: Location-allocation
uid: location-allocation
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

Location-allocation is the process of finding a set of facilities that will best serve demand from surrounding areas. It is a two-fold problem that simultaneously locates facilities and allocates demand points to them. Location-allocation can answer questions such as which site for a new fire station would provide the best response times for a community, which stores a retail company should close to maximize overall demand, or where a factory should be built to minimize the distance to distribution centers.

Location-allocation only supports asynchronous job requests. `location_allocation_job()` takes `facilities`, the candidate locations, and `demand_points`, the locations that generate demand for those facilities.

## Step 1. Authenticate

```{r}
library(sf)
library(arcgisutils)
library(arcgisrouting)

set_arc_token(auth_user())
```

## Step 2. Define facilities and demand points

Each household in `demand_points` has a `weight` representing the size of its demand, and an optional `group_name` for grouping related demand points.

```{r}
facilities <- st_sf(
  name = c("Facility A", "Facility B"),
  facility_type = c(0L, 0L),
  curb_approach = c(0L, 0L),
  geometry = st_sfc(
    st_point(c(-58.557329417999938, -34.587693706999971)),
    st_point(c(-58.460247408999976, -34.683348039999942)),
    crs = 4326
  )
)

demand_points <- st_sf(
  name = c("Household 4", "Household 3", "Household 2", "Household 1"),
  group_name = c("A", "A", NA, NA),
  weight = c(2, 2, 3, 5),
  curb_approach = c(0L, 0L, 0L, 1L),
  geometry = st_sfc(
    st_point(c(-58.664405163999959, -34.614819562999969)),
    st_point(c(-58.514499119999982, -34.496322404999944)),
    st_point(c(-58.54162497599998, -34.788996107999935)),
    st_point(c(-58.40599569799997, -34.637662387999967)),
    crs = 4326
  )
)
```

## Step 3. Submit the job

`location_allocation_job()` creates a geoprocessing job. Call `$start()` to submit it, then `$await()` to poll until it completes and bring the results into memory.

```{r}
job <- location_allocation_job(facilities, demand_points)
job$start()
result <- job$await()
```

`result` is a named list. `solve_succeeded` tells you whether the solve completed, `allocation_lines` shows which facility each demand point was allocated to, and `facilities` and `demand_points` return the input features with the allocation attributes attached.

```{r}
result$solve_succeeded
result$allocation_lines
```

## Choosing a problem type

`problem_type` controls the objective of the analysis. To choose a new restaurant location that captures as many nearby customers as possible, for example, use `problem_type = "maximize_attendance"` together with `number_of_facilities_to_find` to control how many new locations to select, and `travel_direction = "demand_to_facility"` since customers travel to the facility.

```{r}
job <- location_allocation_job(
  facilities,
  demand_points,
  problem_type = "maximize_attendance",
  number_of_facilities_to_find = 2,
  travel_direction = "demand_to_facility"
)

job$start()
job$await()
```
