---
title: Vehicle routing problem
uid: vehicle-routing
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

The vehicle routing problem (VRP) is the process of determining the most effective routes for a fleet of vehicles that need to visit a set of locations, minimizing the overall transportation cost. A furniture store might deliver goods from a central warehouse, a specialized recycling company might route trucks to pick up materials from multiple sites, or a health department might schedule daily inspection visits for its inspectors.

To find an optimized route for a single vehicle rather than a fleet, see [Find routes and directions](/docs/routing/finding-routes.qmd).

`route_vehicles()` takes three inputs: `orders`, the locations to visit, `depots`, the start and return locations for vehicles, and `routes`, the characteristics and constraints of each vehicle.

## Step 1. Authenticate

```{r}
library(sf)
library(arcgisutils)
library(arcgisrouting)

set_arc_token(auth_user())
```

## Step 2. Define orders, depots, and routes

Two vehicles depart from the same depot to fulfill two orders.

```{r}
orders <- st_sf(
  name = c("Order 1", "Order 2"),
  geometry = st_sfc(
    st_point(c(-0.1891, 51.5254)),
    st_point(c(-0.1744, 51.5353)),
    crs = 4326
  )
)

depots <- st_sf(
  name = "Depot1",
  geometry = st_sfc(st_point(c(-0.2, 51.5)), crs = 4326)
)
```

`routes` is a `data.frame` describing each vehicle. Each route names its start and end depot and, optionally, constraints such as capacity.

```{r}
routes <- data.frame(
  name = "Truck1",
  start_depot_name = "Depot1",
  end_depot_name = "Depot1",
  capacities = "40000"
)
```

## Step 3. Solve the vehicle routing problem

```{r}
result <- route_vehicles(orders, depots, routes)
result$routes
```

`route_vehicles()` returns a named list. `solve_succeeded` tells you whether the solve completed, `unassigned_stops` lists any orders that could not be assigned to a route, and `stops` shows the route assignment for each order that was assigned.

```{r}
result$solve_succeeded
result$unassigned_stops
result$stops
```

`usage_cost` and `messages` report the credits consumed and any status or warning messages from the service.

## Generating directions

Set `populate_directions = TRUE` to generate turn-by-turn directions for each route.

```{r}
result <- route_vehicles(
  orders,
  depots,
  routes,
  populate_directions = TRUE
)

result$directions
```

## Larger vehicle routing jobs

`route_vehicles()` makes a direct request, which suits shorter transactions that complete in under five minutes and optimize up to 2 routes and 100 orders. For longer transactions, up to 100 routes and 2,000 orders completing in under four hours, use `route_vehicles_job()` instead, which creates a geoprocessing job.
