---
title: Storing credentials
uid: storing-credentials
---

Authentication requires that you provide sensitive data to the R authentication R functions. These are used to verify your identity to the services that you are calling. These sensitive variables **should never be written down in your code**. 

The packages in the R-ArcGIS Bridge rely heavily on **environment variables**. Environment variables are dynamic values that can be set to store sensitive or configuration information, allowing you to manage and control access to crucial data across various tools and applications, including R scripts.


## Environment variables 

The package `{arcgisutils}` is responsible for handling authorization for all of the R packages. With it, there are a number of different mechanisms for authorizing each requiring a different combination of environment variables. 

Environment variables are fetched using `Sys.getenv()`. They are key-value pairs. For example, to find the path of R run:

```{r}
Sys.getenv("R_HOME")
```

Or to set a value use `Sys.setenv()`:

```{r}
Sys.setenv("ANSWER_TO_EVERYTHING" = 42)
Sys.getenv("ANSWER_TO_EVERYTHING")
```


## Using an `.Renviron` file

Environment variables should _**never**_ be included in your code.
We recommend using an `.Renviron` file at minimum to store your credentials.

You can use the [`usethis`](https://usethis.r-lib.org/) package to edit the file. Ensure it is installed and run the following in your text editor:

:::{.callout-warning}
If you modify environment variables you will need to restart your R session for the change to be registered.
:::

```r
usethis::edit_r_environ()
```
:::{.callout-tip title="`.Renviron` scopes" collapse="true"}
`.Renviron` files can be created at a user level or a project level. It is recommended to use project level `.Renviron` files when deploying scripts to production. If most of your work is interactive then a user level file is sufficient. 

**Project scoped** `.Renviron` files store environment variables that are available for the project _only_. The `.Renviron` is stored at the root folder level.

**User scoped** `.Renviron` files store environment variables in the user's home directory. The environment variables will be available to you in any project you open. However, if you share your project, the environment variables will not be shared with it.
:::

This will open your `.Renviron` file for you to edit. The following are the environment variables used by `{arcgisutils}`.

```bash
# used for OAuth Code & Client
ARCGIS_CLIENT=your-client-id

# used for OAuth Client flow
ARCGIS_SECRET=your-super-secret-key

# used for publishing and Username/Password auth
ARCGIS_USER=your-user-name

# used for API Key auth
ARCGIS_API_KEY=your-developer-api-key

# specify if not using ArcGIS Online
ARCGIS_HOST=https://your-portal.com/ 
```
