---
title: "Using Geoprocessing Services"
uid: using-gp-services

---

Geoprocessing (GP) is one of the fundamental frameworks that powers the ArcGIS System. Many of the tools and APIs available in ArcGIS Online, ArcGIS Enterprise, and ArcGIS Location Platform are exposed using the [Geoprocessing Service Framework](https://developers.arcgis.com/rest/services-reference/enterprise/gp-overview/).

Geoprocessing services are used as the backbone of analysis services, routing, and can even be used to deploy custom tools in a scalable and asynchronous manner.

`{arcgisutils}` provides a structured and unified interface to interacting with GP Services directly from R. 


## Geoprocessing Jobs

We interact with geoprocessing services through a [GP job](https://developers.arcgis.com/rest/services-reference/enterprise/gp-job/) represented by the `arc_gp_job` R6 class. A new geoprocessing job can be created with `arc_gp_job$new()` with the following structure: 

```{r, message=FALSE}
library(arcgisutils)

arc_gp_job$new(
  base_url = "https://your-service-url.com",
  params = list(),
  result_fn = NULL
)
```

::: callout-tip
## Custom packages

If you are creating a new package or wrapper for a specific service it is recommended to create a new class for it with `R6::R6Class("myclass", inherit = arc_gp_job)`.

:::

An `arc_gp_job` has the following methods:

- `new()`: creates a new job from the given url and parameters
- `start()`: starts the job by submitting to the [`/submitJob`](https://developers.arcgis.com/rest/services-reference/enterprise/submit-gp-job/) endpoint
- `cancel()`: cancels the running job by submitting to the [`/cancel`](https://developers.arcgis.com/rest/services-reference/enterprise/cancel-gp-job/) endpoint

In addition to the above methods, the following active bindings are provided: 

- `params`: returns the input parameters as an `arc_form_params` S7 object (access the underlying list with `job$params@params`)
- `status`: returns the status of the job as an `arc_job_status` S7 object 
- `results`: fetches the results of the job and applies the `result_fn` to the results by calling the [`/results`](https://developers.arcgis.com/rest/services-reference/enterprise/results/) endpoint


## Parameters


The geoprocessing framework has [service data types](https://developers.arcgis.com/rest/services-reference/enterprise/gp-data-types/) that can be provided to a geoprocessing job. Most geoprocessing parameters can be created and parsed using arcgisutils. 

| Parameter Type | Write | Read |
|---|---|---|
| GPFeatureRecordSetLayer | `as_esri_featureset()` | `parse_gp_feature_record_set()` |
| GPRecordSet | `as_record_set()` | `parse_gp_record_set()` |
| GPRasterDataLayer | `as_gp_raster_layer()` | `yyjsonr::read_json_str()` |
| Field | `as_fields()` creates a list of fields. Convert to json with  `yyjsonr::write_json_str(x, auto_unbox = TRUE)`.| `yyjsonr::read_json_str()` |
| GPBoolean, GPDouble, GPLong, GPString | `yyjsonr::write_json_str(x, auto_unbox = TRUE)` | `yyjsonr::read_json_str()` |
| GPLinearUnit | `as_gp_linear_unit()` | `parse_gp_linear_unit()` |
| GPArealUnit | `as_gp_areal_unit()` | `parse_gp_areal_unit()` |
| GPDate | `as_gp_date()` | `parse_gp_date()` |
| GPSpatialReference | `as_spatial_reference()` | `parse_spatial_reference()` |

Many geoprocessing services require a **feature set** which can be created with `as_esri_featureset()`.

### Unsupported parameter types ❌

The following parameter types are not yet supported. 

- `GPValueTable`
- `GPComposite`
- `GPMultiValue`
- `GPDataFile`
- `GPEnvelope`
- `GPCoordinateSystem`
- `GPStringHidden`
- `GPSQLExpression`

::: callout-note

Please [create an issue](https://github.com/R-ArcGIS/arcgisutils/issues/new) if you have a use case for an unsupported parameter type.

:::

## Processing Outputs

An `arc_gp_job` has a `result_fn` which is applied to the raw json string that that is returned from the `$result` active binding. 

In many cases, the result from a GP service is a feature set output. To convert that into an `{sf}` object use the `parse_gp_feature_record_set()` function. Otherwise, the result is parsed using `RcppSimdJson::fparse()` to convert the result into a list. 


## Example: Trace Downstream service

As a complete example we will call the [Trace Downstream](https://developers.arcgis.com/rest/elevation-analysis/trace-downstream/) elevation service. This example will return a linestring of the downstream flowpath from the provided point. 

Here we create a single point as an `sfc` object which will be provided to the `InputPoints` parameter as a FeatureSet. 

```{r message=FALSE}
library(sf)

# create input points
input_points <- st_sfc(
  st_point(c(-159.548936, 21.955888)),
  crs = 4326
)
```



```{r}
library(arcgisutils)

# authenticate
set_arc_token(auth_user())

# store the service url
service_url <- "https://hydro.arcgis.com/arcgis/rest/services/Tools/Hydrology/GPServer/TraceDownstream"

# create a new job
downstream_job <- arc_gp_job$new(
  base_url = service_url,
  params = list(InputPoints = as_esri_featureset(input_points), f = "json"),
  # the result is a GPRecordFeatureSet
  result_fn = parse_gp_feature_record_set,
  token = arc_token(),
)

# print the job
downstream_job
```


This creates a new `arc_gp_job` object. The object stores the logic of the job but it has not yet been instantiated. To start the job we need to explicitly call `$start()`.

```{r}
downstream_job$start()
```

This starts the job asynchronously. You can let this job run to completion without having to worry about it. Or, if you want to wait for it to complete and bring the result in to memory, use the `$await()` method on the job.

```{r}
downstream_job$await()
```

This returns a `GPFeatureRecordSetLayer` result which is a list containing the parameter name, data type, and the geometry associated with the result. 

