---
title: Overview
uid: layers-overview
---

Layers represent represent a data service hosted by ArcGIS Online, ArcGIS Enterprise, Platform, or even an ArcGIS Hub site. Layers are how you will interact with your remotely hosted data. 

The `{arcgis}` R metapackage is able to work with the most common types of data services: feature services and image servers. The types of layers that the R-ArcGIS Bridge does not yet support are vector and map tile services. 


# Types of services 

When working with these data services, each service type is assigned their own S3 class. 


| Class | Description |
| ----- | ----------- |
| `FeatureLayer` | Represents vector data. Can contain Point, MultiPoint, Polyline, or Polygon data | 
| `Table` | A type of FeatureLayer in which there is no geometry present. Represents tabular data. | 
| `ImageServer` | Provides access to raster imagery.  | 
| `FeatureServer` | A collection of FeatureLayer and Tables. | 
| `GroupLayer` | A collection of FeatureLayer and Tables. | 
| `MapServer` | A collection arbitrary data service types. | 

# Accessing data services

The R package `{arcgislayers}` provides access to services. Working with data services follows a pattern:

- first, create a reference to a remote data service
- query the data service and bring the results back into R

## Creating data service objects

Using `{arcgislayers}`, the function `arc_open()` will create a reference to a service based on its url. 

```{r}
library(arcgis)

# feature service url
furl <- "https://services3.arcgis.com/ZvidGQkLaDJxRSJ2/arcgis/rest/services/PLACES_LocalData_for_BetterHealth/FeatureServer/4"

# feature layer
flayer <- arc_open(furl)
flayer
```

## Querying feature services 

To bring data from the service into memory in R, you must use `arc_select()`. 


:::{.callout-tip}
Always bring as little data as possible into memory. Utilize the `fields` and `where` arguments. 
:::

```{r message=FALSE}
arc_select(
  flayer,
  where = "TotalPopulation > 100000",
  fields = c("objectid", "totalpopulation", "zcta5")
)
```

